### 2018-04-04 chapter6/ms.tex
Redo Ch6. Start Intro

%UNLIKE IN A REGULAR TEX FILE, DON'T PUT ANY PREAMBLE MATERIAL HERE
\section*{Abstract}
We present preliminary studies into the development of efficient and effective strategies for the usage of LSST in optical follow-up of gravitational wave events. Specifically, we focus on the detection and characterization of kilonovae across a wide range of ejecta properties and composition. We show that the main LSST survey will achieve an average efficiency for kilonova detection of \apx3\%, with only a handful of sources characterized well enough for detailed study. In contrast, triggered target-of-opportunity follow up with a time investment of just 1 hour per event will be able to easily detect and characterize the brightest kilonova models, provided observations commence within \apx9 hours of the GW trigger. This suggests that LSST has the opportunity to play an important role in the growth of multi-messenger astronomy by facilitating highly efficient kilonova searches.

@@

Here we expand on the groundwork established by \citet{Scolnic+18} along two new avenues. First, we expand the range of kilonovae models considered in the LSST main survey. This is accomplished by considering a wider range of ejecta masses and composition to fully explore the potential range of kilonovae brightness and timescales. Second, we expand beyond the main survey cadence and explore the detectability of kilonovae in triggered target-of-opportunity follow-up observations. We explore not only the ability of such observations to detect kilonovae, but also identify and characterize their behavior.

@@

In this work, we consider the detectability and characterization of kilonovae in two unique observing situations. The first is the serendipitous detection of kilonovae in the LSST main survey independent of a GW trigger from LIGO. The second is triggered target-of-opportunity observations conducted by LSST in response to a GW trigger. We generate simulated observations for both scenarios as follows:

\subsection{Main Cadence Simulations With OpSim}
\label{sec:ch6_opsim}
We produce simulated observations of kilonovae in the LSST main cadence using the LSST Operations Simulator \citep[OpSim,][]{OpSim1,OpSim2}. This is a publicly available package designed to produce realistic simulations of LSST scheduling and imaging over the ten year duration of the survey. These simulations provide realistic information about observations including cadence and science program requirements, observing conditions due to telescope or environmental factors, and image characteristics. We use the most recent OpSim reference run ({\tt minion\_1016}) to produce a catalog of observations including times, sky position, photometric errors, and limiting magnitudes.

@@

\subsection{Target of Opportunity Simulations With SNANA}
\label{sec:ch6_snana}
We parametrize our triggered target-of-opportunity (ToO) strategies based on the expected time investment per night and total number of nights spent observing. In this work we assume that the typical LIGO localization region is \apx10~deg$^2$ and can therefore be observed in just two LSST pointings, which are necessary to cover chip gaps in the CCD mosaic. We assume that each pointing consists of $grizY$ observations lasting 30~s each with per filter overheads of 90~s. This puts the total time investment at 10 minutes per pointing, or 20 minutes per night. We assume 7 nights of observing for a total telescope time investment of \apx2.5~hours. To test the effect of rapid cadence observations, we include a second epoch on the first night, taken 3 hours after the initial epoch.

We simulate these ToO observations using the SNANA public software package \citep{SNANA}. SNANA is a supernova analysis and simulation package designed to produce realistic supernovae light curves for the optimization of time-domain surveys and extraction of cosmological parameters. SNANA generates these light curves using a model transient SED, realistic filter transmission functions, and a database of observation dates and conditions. We run SNANA using our kilonova model SEDs and LSST filter transmission functions, and generate the database of observing conditions by pulling key information from the OpSim simulations (e.g., sky noise, PSF, observed zero point, see \cref{sec:ch6_opsim}). We then mutate the observing dates to match those required for our ToO cadence described above.

We inject kilonova following the same procedure outlined in \cref{sec:ch6_opsim} with several key differences. First, we do not inject sources at randon sky locations, but instead at the location LSST is observing in the SNANA database. Second, in order to probe a realistic range of response times for the ToO trigger, we inject kilonova at a uniformly sampled time from 3--24 hours prior to the first epoch. Lastly, SNANA randomly selects an injection distance using a volume-weighted scheme to generate a realistic distribution. We therefore, inject 10,000 kilonovae across the entire volume ($D_L \apx450$~Mpc) to ensure that we have enough nearby sources. Since the distance injection in SNANA is based on redshift, a handful of sources end up with a distance greater than 450~Mpc. To maintain consistency with our OpSim volume, we truncate these sources leaving a final sample of 9171 light curves for our analysis in \cref{sec:ch6_snana_results}.

\section{Analysis}
\label{sec:ch6_analysis}
The analysis of the simulated light curves presented here is focused along two major lines of thought. In the context of kilonovae detected in the LSST main survey, we are primarily concerned with not only how many kilonovae will be detected, but also how many of those detections will yield enough light curve information to make statements about the ejecta (e.g., mass, velocity, composition). In the context of targeted ToO observations, we are not only concerned with these issues of detection and characterization, but also the impact of program design (e.g., telescope time investment) on the science returns.

@@

\begin{figure}[!t]
\begin{center}
\hspace*{-0.1in}
\scalebox{1.}
{\includegraphics[width=0.75\textwidth]{./figs/chapter6/f1.pdf}}
\caption{\singlespace Histograms showing the number of detections as a function of model and distance, for those models that are detected. We note that as a result of the cadence of the LSST main survey we are not always able to obtain crucial information about the rise and color.}
\label{fig:ch6_eff_hist}
\end{center}
\end{figure}

@@

\begin{figure}[!t]
\begin{center}
\hspace*{-0.1in}
\scalebox{1.}
{\includegraphics[width=0.725\textwidth]{./figs/chapter6/f2.pdf}}
\caption{\singlespace Histogram of events as a function of time to first or last detection. As with \Cref{fig:ch6_eff_hist}, we only plot those models that are detected. The LSST main cadence is only able to track a limited number of events out to longer times than \apx~a few days. This combined with the poor rise and color constraints shown in \Cref{fig:ch6_eff_hist} means that kilonovae detected in the LSST main survey will be difficult to detect and characterize.}
\label{fig:ch6_t_hist}
\end{center}
\end{figure}

@@

\begin{figure}[!t]
\begin{center}
\hspace*{-0.1in}
\scalebox{1.}
{\includegraphics[width=0.9\textwidth]{./figs/chapter6/f3.pdf}}
\caption{\singlespace Detection efficiency as a function of distance for each of the detected models. As expected, the efficiency is high (\apx45\%) for nearby sources ($D_L \apx50$~Mpc), but rapidly tapers off to \apx15\% at the edge of the LIGO BNS detection range $(D_L \gtrsim 400$~Mpc).}
\label{fig:ch6_dist_eff}
\end{center}
\end{figure}

@@

\begin{figure}[!t]
\begin{center}
\hspace*{-0.1in}
\scalebox{1.}
{\includegraphics[width=0.9\textwidth]{./figs/chapter6/f5.pdf}}
\caption{\singlespace Efficiency of our ToO searches as a function of distance. The numbers plotted here are relative to the total number of injected sources (i.e., including models not detected). The {\it per model} detection efficiency is 100\% for the bright KN models plotted. The efficiency curve is flat as a function of distance indicating that ToO efforts with LSST can be effective out to the maximal LIGO BNS detection distance.}
\label{fig:ch6_snana_dist_eff}
\end{center}
\end{figure}

@@

\begin{figure}[!t]
\begin{center}
\hspace*{-0.1in}
\scalebox{1.}
{\includegraphics[width=0.9\textwidth]{./figs/chapter6/f6.pdf}}
\caption{\singlespace Efficiency of our ToO searches to detect brightening in the kilonova light curve in at least one filter as a function of starting time. The efficiency here is computed relative to the number of detected sources in a given time bin $(\Delta_t \apx3$~hrs). There is a sharp drop off in efficiency after just a few hours. This efficiency drops to 0 under all observing scenarios when just a single epoch is obtained on the first night.}
\label{fig:ch6_snana_t_eff}
\end{center}
\end{figure}

>>>>

%UNLIKE IN A REGULAR TEX FILE, DON'T PUT ANY PREAMBLE MATERIAL HERE
\section*{Abstract}
We present preliminary studies into the development of efficient and effective strategies for the usage of LSST in optical follow-up of gravitational wave events. Specifically, we focus on the detection and characterization of kilonovae across a wide range of ejecta properties and composition. We show that the main LSST survey will achieve an average efficiency for kilonova detection of \apx3\%, with only a handful of sources characterized well enough for detailed study. In contrast, triggered target-of-opportunity follow up with a time investment of just 1 hour per event will be able to detect and characterize the brightest kilonova models, provided observations commence promptly after the GW trigger. This suggests that LSST has the opportunity to play an important role in the growth of multi-messenger astronomy by facilitating highly efficient kilonova searches.

@@

Here we expand on the groundwork established by \citet{Scolnic+18} along two new avenues. First, we expand the range of kilonovae models considered in the LSST main survey. This is accomplished by considering a wider range of ejecta masses and composition to fully explore the potential range of kilonovae brightness and timescales. Second, we expand beyond the main survey cadence to discuss the benefits of targeted target-of-opportunity (ToO) observations when compared to the main survey.

%and explore the detectability of kilonovae in triggered target-of-opportunity follow-up observations. We explore not only the ability of such observations to detect kilonovae, but also identify and characterize their behavior.

@@

%In this work, we consider the detectability and characterization of kilonovae in two unique observing situations. The first is the serendipitous detection of kilonovae in the LSST main survey independent of a GW trigger from LIGO. The second is triggered target-of-opportunity observations conducted by LSST in response to a GW trigger. We generate simulated observations for both scenarios as follows: 
%
%\subsection{Main Cadence Simulations With OpSim}
%\label{sec:ch6_opsim}

We produce simulated observations of kilonovae in the LSST main cadence using the LSST Operations Simulator \citep[OpSim,][]{OpSim1,OpSim2}. This is a publicly available package designed to produce realistic simulations of LSST scheduling and imaging over the ten year duration of the survey. These simulations provide realistic information about observations including cadence and science program requirements, observing conditions due to telescope or environmental factors, and image characteristics. We use the most recent OpSim reference run ({\tt minion\_1016}) to produce a catalog of observations including times, sky position, photometric errors, and limiting magnitudes.

@@

%\subsection{Target of Opportunity Simulations With SNANA}
%\label{sec:ch6_snana}
%We parametrize our triggered target-of-opportunity (ToO) strategies based on the expected time investment per night and total number of nights spent observing. In this work we assume that the typical LIGO localization region is \apx10~deg$^2$ and can therefore be observed in just two LSST pointings, which are necessary to cover chip gaps in the CCD mosaic. We assume that each pointing consists of $grizY$ observations lasting 30~s each with per filter overheads of 90~s. This puts the total time investment at 10 minutes per pointing, or 20 minutes per night. We assume 7 nights of observing for a total telescope time investment of \apx2.5~hours. To test the effect of rapid cadence observations, we include a second epoch on the first night, taken 3 hours after the initial epoch.
%
%We simulate these ToO observations using the SNANA public software package \citep{SNANA}. SNANA is a supernova analysis and simulation package designed to produce realistic supernovae light curves for the optimization of time-domain surveys and extraction of cosmological parameters. SNANA generates these light curves using a model transient SED, realistic filter transmission functions, and a database of observation dates and conditions. We run SNANA using our kilonova model SEDs and LSST filter transmission functions, and generate the database of observing conditions by pulling key information from the OpSim simulations (e.g., sky noise, PSF, observed zero point, see \cref{sec:ch6_opsim}). We then mutate the observing dates to match those required for our ToO cadence described above.
%
%We inject kilonova following the same procedure outlined in \cref{sec:ch6_opsim} with several key differences. First, we do not inject sources at randon sky locations, but instead at the location LSST is observing in the SNANA database. Second, in order to probe a realistic range of response times for the ToO trigger, we inject kilonova at a uniformly sampled time from 3--24 hours prior to the first epoch. Lastly, SNANA randomly selects an injection distance using a volume-weighted scheme to generate a realistic distribution. We therefore, inject 10,000 kilonovae across the entire volume ($D_L \apx450$~Mpc) to ensure that we have enough nearby sources. Since the distance injection in SNANA is based on redshift, a handful of sources end up with a distance greater than 450~Mpc. To maintain consistency with our OpSim volume, we truncate these sources leaving a final sample of 9171 light curves for our analysis in \cref{sec:ch6_snana_results}.

\section{Analysis}
\label{sec:ch6_analysis}
The analysis of the simulated light curves presented here is focused along two major lines of thought. In the context of kilonovae detected in the LSST main survey, we are primarily concerned with not only how many kilonovae will be detected, but also how many of those detections will yield enough light curve information to make statements about the ejecta (e.g., mass, velocity, composition). We then wish to compare these results to qualitative expectations for a targeted ToO program.

%In the context of targeted ToO observations, we are not only concerned with these issues of detection and characterization, but also the impact of program design (e.g., telescope time investment) on the science returns.

@@

%\begin{figure}[!t]
%\begin{center}
%\hspace*{-0.1in}
%\scalebox{1.}
%{\includegraphics[width=0.75\textwidth]{./figs/chapter6/f1.pdf}}
%\caption{\singlespace Histograms showing the number of detections as a function of model and distance, for those models that are detected. We note that as a result of the cadence of the LSST main survey we are not always able to obtain crucial information about the rise and color.}
%\label{fig:ch6_eff_hist}
%\end{center}
%\end{figure}

@@

%\begin{figure}[!t]
%\begin{center}
%\hspace*{-0.1in}
%\scalebox{1.}
%{\includegraphics[width=0.725\textwidth]{./figs/chapter6/f2.pdf}}
%\caption{\singlespace Histogram of events as a function of time to first or last detection. As with \Cref{fig:ch6_eff_hist}, we only plot those models that are detected. The LSST main cadence is only able to track a limited number of events out to longer times than \apx~a few days. This combined with the poor rise and color constraints shown in \Cref{fig:ch6_eff_hist} means that kilonovae detected in the LSST main survey will be difficult to detect and characterize.}
%\label{fig:ch6_t_hist}
%\end{center}
%\end{figure}

@@

%\begin{figure}[!t]
%\begin{center}
%\hspace*{-0.1in}
%\scalebox{1.}
%{\includegraphics[width=0.9\textwidth]{./figs/chapter6/f3.pdf}}
%\caption{\singlespace Detection efficiency as a function of distance for each of the detected models. As expected, the efficiency is high (\apx45\%) for nearby sources ($D_L \apx50$~Mpc), but rapidly tapers off to \apx15\% at the edge of the LIGO BNS detection range $(D_L \gtrsim 400$~Mpc).}
%\label{fig:ch6_dist_eff}
%\end{center}
%\end{figure}

@@

%\begin{figure}[!t]
%\begin{center}
%\hspace*{-0.1in}
%\scalebox{1.}
%{\includegraphics[width=0.9\textwidth]{./figs/chapter6/f5.pdf}}
%\caption{\singlespace Efficiency of our ToO searches as a function of distance. The numbers plotted here are relative to the total number of injected sources (i.e., including models not detected). The {\it per model} detection efficiency is 100\% for the bright KN models plotted. The efficiency curve is flat as a function of distance indicating that ToO efforts with LSST can be effective out to the maximal LIGO BNS detection distance.}
%\label{fig:ch6_snana_dist_eff}
%\end{center}
%\end{figure}

@@

%\begin{figure}[!t]
%\begin{center}
%\hspace*{-0.1in}
%\scalebox{1.}
%{\includegraphics[width=0.9\textwidth]{./figs/chapter6/f6.pdf}}
%\caption{\singlespace Efficiency of our ToO searches to detect brightening in the kilonova light curve in at least one filter as a function of starting time. The efficiency here is computed relative to the number of detected sources in a given time bin $(\Delta_t \apx3$~hrs). There is a sharp drop off in efficiency after just a few hours. This efficiency drops to 0 under all observing scenarios when just a single epoch is obtained on the first night.}
%\label{fig:ch6_snana_t_eff}
%\end{center}
%\end{figure}
